---
title: "Untitled"
author: "Juliano Palacios Abrantes"
date: "2024-04-09"
output: 
  html_document: 
    toc: yes
    keep_md: yes
---

# Script objective

This script extract the different data sources for straddling stocks identification and shifts for the West Africa region. In short this script follows these 4 steps:

1.- Define spatial boundaries for "West Africa"

2.- Extract transboundary species for the region from [Palacios-Abrantes et al., (2020)](https://www.nature.com/articles/s41598-020-74644-2)

3.- Extract shifts in these species' transboundary stocks according to [Palacios-Abrantes et al., 2022](https://onlinelibrary.wiley.com/doi/10.1111/gcb.16058)

4.- Extract shifts in these species' straddling stocks according to [Palacios-Abrantes et al., in prep]()


```{r setup, eval = T, echo = F, warning = F, message = F, results = 'hide'}

library(MyFunctions)

#### Project's Library
packages <- c(
  "tidyverse",
  "here", # for dbem_import `here()`da
  "data.table", #dbem_import `fread()`
  "readxl", # for reading excell files
  "janitor", # for clearing names
  # "geosphere", # estimate distances between points `distm()`
  # "ggrepel",
  # "zoo", # for average mean
  # "parallel", # for mclapply,
  "sf", # for mapping
  "st", # for mapping
  # "rgdal", #Spatial analysis
  # "tools", #Spatial analysis 
  # "zeallot", # for Juanito's map
  # "gmt", # for estimating distances between points
  "viridis"
  # "cowplot",
  # "rfishbase",
  # "stargazer", # For nice lm tables
  # "tidytext", # for boxplot order
  # "moments", # for skewness and kurto
  # "pgirmess" # for non parametric post hoc test
)

my_lib(packages)

# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

```

# 1.- Set spatial boundaries

## NOTES

- Western Sahara region is allocated to Morocco's as "South EEZ"
- Do we include the Canary Islands (Spain), Madeira Isl. (Portugal), and Cape Verde?

```{r spatial_boundaries}

# Load SAU spatial info
sau_info <- my_path("Spa","SAU",name = "sau_matching_names.csv", read = T)

# West african countries
nwa_eez <- c("Senegal","Mauritania","Western Sahara","Morocco (South)","Morocco (Central)","Canary Isl. (Spain)","Madeira Isl. (Portugal)","Cape Verde","Gambia")

# Load SAU sf
sau_eez <- my_sf("SAU")

sau_eez %>% 
  filter(name %in% nwa_eez) %>% 
  ggplot() +
  geom_sf(
    aes(
      color = name,
      fill = name
      )
  ) +
  my_land_map() +
  coord_sf(
    xlim = c(-30,-5),
    ylim = c(10,40)
  )

```


## 2.- Extract transboundary species for the region

```{r}

# Read species taxon info
taxon_inf <- my_data("sau_species")

# Read transboundary species database from Palacios-Abrantes et al 2020 (FishForVisa)
Transboundary_spp <- my_path("G", extra_path = "FishForVisa/Results/", name="Transboundary_spp.csv", read = TRUE)


nwa_trans_spp <- Transboundary_spp %>% 
  filter(eez_name %in% nwa_eez) %>% 
  left_join(taxon_inf) %>% 
  select(taxon_key,
         taxon_name,
         common_name,
         eez_name,
         eez_neighbour)

# Double check all eezs are there, yes.
# unique(nwa_trans_spp$eez_name)

write_csv(nwa_trans_spp,
          my_path("D","species","nwa_trans_spp.csv"))

```

### Summary of results

There are 149 species identified as transboundary in the region: 

```{r}
nwa_trans_spp %>% 
  group_by(common_name,taxon_name) %>% 
  tally() %>% 
    # View()
  knitr::kable(
    caption = "Transboundary species identified for the region. n = the number of times the species is shared (i.e., a proxy of stock)"
               )
```

## 3.- Extract shifts in these species' transboundary stocks

```{r}

# Load SSR data
ssr_data_path <- my_path("G","EmergingFish/Results/Proportion_2005c")

# Get the sepcies we want 
ssr_to_read <- paste0(ssr_data_path,"proportion_",unique(nwa_trans_spp$taxon_key),".csv")


nwa_ssr_data <- bind_rows(
    lapply(ssr_to_read, read_proportion, eez = nwa_eez)
  )

# Quick check data
length(unique(nwa_ssr_data$taxon_key)) # 147 taxon

unique(nwa_ssr_data$eez_name) # 8

# Save data for futher analysis
write_csv(nwa_ssr_data,
          my_path("D","species","nwa_ssr_data.csv")
          )

```


