---
title: "Untitled"
author: "Juliano Palacios Abrantes"
date: "2024-04-09"
output: 
  pdf_document: 
    toc: yes
    keep_md: yes
---

# Script objective

This script extract the different data sources for straddling stocks identification and shifts for the West Africa region. In short this script follows these 4 steps:

1.- Define spatial boundaries for "West Africa"

2.- Extract transboundary species for the region from [Palacios-Abrantes et al., (2020)](https://www.nature.com/articles/s41598-020-74644-2)

3.- Extract shifts in these species' transboundary stocks according to [Palacios-Abrantes et al., 2022](https://onlinelibrary.wiley.com/doi/10.1111/gcb.16058)

4.- Extract shifts in these species' straddling stocks according to [Palacios-Abrantes et al., in prep]()


```{r setup, eval = T, echo = F, warning = F, message = F, results = 'hide'}

library(MyFunctions)

#### Project's Library
packages <- c(
  "tidyverse",
  "here", # for dbem_import `here()`da
  "data.table", #dbem_import `fread()`
  "readxl", # for reading excell files
  "janitor", # for clearing names
  # "geosphere", # estimate distances between points `distm()`
  # "ggrepel",
  # "zoo", # for average mean
  # "parallel", # for mclapply,
  "sf", # for mapping
  "st", # for mapping
  # "rgdal", #Spatial analysis
  # "tools", #Spatial analysis 
  # "zeallot", # for Juanito's map
  # "gmt", # for estimating distances between points
  "viridis"
  # "cowplot",
  # "rfishbase",
  # "stargazer", # For nice lm tables
  # "tidytext", # for boxplot order
  # "moments", # for skewness and kurto
  # "pgirmess" # for non parametric post hoc test
)

my_lib(packages)

# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

```

# 1.- Set spatial boundaries

## NOTES

- Western Sahara region is allocated to Morocco's as "South EEZ"
- Do we include the Canary Islands (Spain), Madeira Isl. (Portugal), and Cape Verde?

```{r spatial_boundaries, eval = T, echo = F, message = F, warning = F,}

# Load SAU spatial info
sau_info <- my_path("Spa","SAU",name = "sau_matching_names.csv", read = T)

# West African countries
nwa_eez <- c("Senegal","Mauritania","Western Sahara","Morocco (South)","Morocco (Central)","Canary Isl. (Spain)","Madeira Isl. (Portugal)","Cape Verde","Gambia")

# Load SAU sf
sau_eez <- suppressMessages(my_sf("SAU")) 

sau_eez %>% 
  filter(name %in% nwa_eez) %>% 
  ggplot() +
  geom_sf(
    aes(
      color = name,
      fill = name
      )
  ) +
  my_land_map() +
  coord_sf(
    xlim = c(-30,-5),
    ylim = c(10,40)
  )

```


## 2.- Extract transboundary species for the region

```{r transboundary_extraction, eval = F, echo = F}

# Read species taxon info
taxon_inf <- my_data("sau_species")

# Read transboundary species database from Palacios-Abrantes et al 2020 (FishForVisa)
Transboundary_spp <- my_path("G", extra_path = "FishForVisa/Results/", name="Transboundary_spp.csv", read = TRUE)


nwa_trans_spp <- Transboundary_spp %>% 
  filter(eez_name %in% nwa_eez) %>% 
  left_join(taxon_inf) %>% 
  select(taxon_key,
         taxon_name,
         common_name,
         eez_name,
         eez_neighbour)

# Double check all eezs are there, yes.
# unique(nwa_trans_spp$eez_name)

write_csv(nwa_trans_spp,
          my_path("D","species","nwa_trans_spp.csv"))

```

### Summary of results

There are 149 species identified as transboundary in the region: 

```{r tran_example, eval = T, echo = F, message= F, error=F}

my_path("D","species","nwa_trans_spp.csv", read = T) %>% 
  group_by(common_name,taxon_name) %>% 
  tally() %>% 
    # View()
  head(10) %>% 
  knitr::kable(
    caption = "Transboundary species identified for the region. n = the number of times the species is shared (i.e., a proxy of stock)"
               )
```

## 3.- Extract shifts in these species' transboundary stocks

```{r get_ssr_trans, eval = F, echo = F}

# West African countries
nwa_eez <- c("Senegal","Mauritania","Western Sahara","Morocco (South)","Morocco (Central)","Canary Isl. (Spain)","Madeira Isl. (Portugal)","Cape Verde","Gambia")

# Get species list
trans_spp <- my_path("D","species","nwa_trans_spp.csv", read = T) %>% 
  pull(taxon_key) %>% 
  unique()

# Run protocol, note that we had to modify the code a bit from the original analysis. See functions folder
suppressWarnings(
  suppressMessages(
    nwa_ssr_trans_data <- 
      bind_rows(
        lapply(trans_spp,summary_trans_ssr, nwa_countries = nwa_eez)
      )%>% 
      # re organize variables
      select(
        taxon_key:eez_neighbour,time_historical,time_early,time_mid
      ) 
  )
) 


# Save data for further analysis
write_csv(nwa_ssr_trans_data,
          my_path("D","species","nwa_ssr_trans_data")
          )

  

```

### Summary of results

```{r summary_trans_ssr, eval = T, echo = F, message = F, warning = F, fig.cap = "Stock Share Ratio (SSR) in percentage (left) between each EEZ of all stocks by time frame (bottom). reading  top to right, for example, Cape Verte has 75% of the SSR of one stock shared with Mauritania who has the 25% (100-75) in the historical time period."}

my_path("D",extra_path = "species", name = "nwa_ssr_trans_data.csv", read = T) %>% 
  # group_by(eez_name,eez_neighbour) %>% 
  # summarise_at(
  #   vars(matches("time_")), ~mean(., na.rm = T)
  # ) %>% 
  gather("time_period","value",time_historical:time_mid) %>% 
  mutate(order = ifelse(time_period == "time_historical","a_hist",
                        ifelse(time_period == "time_early","b_early","c_mid"))
         ) %>% 
  arrange(eez_name) %>% 
  slice(1:936) %>% 
  # View()
  ggplot() +
    geom_boxplot(
      aes(
        x = order,
        y = value
        # fill = order,
        # color = order
      ),
      # stat = "identity",
      # position = "dodge"
  ) +
  ylab("Stock Share Ratio across all stocks (%)") +
  xlab("Time period") +
  # coord_flip() +
  scale_fill_viridis_d("Time period") +
  scale_color_viridis_d("Time period") +
  # facet_wrap(eez_neighbour~eez_name)
  facet_grid(eez_neighbour~eez_name) +
  theme_bw()
```

## 3.- Extract shifts in these species' straddling stocks

```{r get_ssr_strad, eval = F, echo = F}

# West African countries
nwa_eez <- c("Senegal","Mauritania","Western Sahara","Morocco (South)","Morocco (Central)","Canary Isl. (Spain)","Madeira Isl. (Portugal)","Cape Verde","Gambia")

# Get straddling species list
stradd_spp <- read_csv("~/GitHub/MigrantFish/results/supplements/straddling_species.csv") %>% 
  select(taxon_name) %>% 
  left_join(my_data("dbem_species")) 

# Save data for further analysis
write_csv(stradd_spp,
          my_path("D","species","nwa_stradd_spp.csv")
          )

stradd_spp <- stradd_spp %>% 
  pull(taxon_key)

# Get realms
realm_to_eez <- read_csv("~/GitHub/MigrantFish/results/supplements/realm_to_eez.csv") %>% 
  clean_names() %>% 
  filter(eez %in% nwa_eez)

realms <- unique(realm_to_eez$realm)

# Load data
stradd_spp_path <- paste0("~/Library/CloudStorage/OneDrive-UBC/Data/MigrantFish/Results/Per_change_realm_rfmo/", stradd_spp,"_perchg.csv")


nwa_ssr_stradd_data <- 
  bind_rows(
    lapply(stradd_spp_path, read_ssr_stradd, realm = realms)
  )

unique(nwa_ssr_stradd_data)

# Save data for further analysis
write_csv(nwa_ssr_stradd_data,
          my_path("D","species","nwa_ssr_stradd_data.csv")
          )


```

### Summary of results


```{r summary_stradd_ssr, eval = T, echo = F, message = F, warning = F, fig.cap = "Number of stocks shifting their Stock Share Ratio (SSR) in percentage (left) between EEZs (grouped by realm) and ICCAT by time period (bottom). Colors represent the direction of change. For example, in all cases most stocks wont shift, but when they do they will likeley go to the high seas. Note that the varioation is given because there are two realms in the region"}

my_path("D","species","nwa_ssr_stradd_data.csv", read = T) %>% 
  group_by(rcp,period,rfmo_name,realm_name,change) %>%
  tally() %>% 
  filter(rcp == "rcp85") %>%
  mutate(direction = ifelse(change == "gain","RFMO",
                            ifelse(change == "lost","EEZs","Same"))
         ) %>% 
  # View()
  ggplot() +
    geom_boxplot(
      aes(
        x = period,
        y = n,
        fill = direction,
        # color = order
      ),
  ) +
  ylab("Number of stocks shifting") +
  # coord_flip() +
  scale_fill_viridis_d("Direction of change") +
  # scale_color_viridis_d("Time period") +
  # facet_wrap(eez_neighbour~eez_name)
  # facet_grid(eez_neighbour~eez_name) +
  theme_bw()

```

